---
title: "Human Activity Recognition"
author: "Andrey Shadrikov"
date: "Sunday, January 25, 2015"
output: html_document
bibliography: pml.bib
---

##Data Overview

Data about human activity [from @ugulino2012wearable] taking different exercising.

Let's load our training data (testing data we leave for prediction only).
```{r,warning=FALSE,message=FALSE}
library(caret)
data <- read.table('pml-training.csv', header=TRUE, sep=',', na.strings=c('NA', '#DIV/0!'))
summary(data)
```

There are lot of NAs in our data (more that $97.5\%$) in particular features.
Given there are $126$ such features and we can reduce their number to $34$ removing them give us not only solution from NAs in data but also the acceleration of optimization in machine learning algorithms.

##Preprocessing

Removing all "bad" features we get new dataset:
```{r}
data.reduced <- data[,c('user_name', 'raw_timestamp_part_1', 'raw_timestamp_part_2', 'new_window', 'num_window', 'roll_belt', 'pitch_belt', 'yaw_belt', 'total_accel_belt', 'gyros_belt_x', 'gyros_belt_y', 'gyros_belt_z', 'accel_belt_x' ,'accel_belt_y' ,'accel_belt_z' ,'magnet_belt_x' ,'magnet_belt_y' ,'magnet_belt_z' ,'roll_arm' ,'pitch_arm' ,'yaw_arm', 'total_accel_arm', 'gyros_arm_x', 'gyros_arm_y', 'gyros_arm_z', 'accel_arm_x', 'accel_arm_y', 'accel_arm_z', 'magnet_arm_x', 'magnet_arm_y', 'magnet_arm_z', 'roll_dumbbell', 'pitch_dumbbell', 'yaw_dumbbell', 'classe')]
```

One can note that there is feature "cvtd_timestamp" that also was removed.
It's because we already has timestamps in good for learning algorithm form.
Thus we can leave this feature behind.

Now we can set our control of training process (using 20-fold cross validation).
```{r}
trainIdx <- createDataPartition(y=data.reduced$classe, p=0.8, list=FALSE)
train <- data.reduced[trainIdx,]
test <- data.reduced[-trainIdx,]
train_control <- trainControl(method="cv", number=10)
```

##Model Fitting

Let's try random forest to fit and understand our data.
```{r, warning=FALSE, message=FALSE}
model <- train(classe~., data=train, trControl=train_control, method="rf")
predictions <- predict(model, test)
confusionMatrix(predictions, test$classe)
```

We build quite good classifier.
Let's use it for prediction.
```{r}
finalModel <- model
```

##Predicting

```{r, echo=FALSE}
pml_write_files <- function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```

Using our final model predicting our goal:
```{r}
data.test <- read.table('pml-testing.csv', header=TRUE, sep=',', na.strings=c('NA', '#DIV/0!'))
data.test.reduced <- data.test[,c('user_name', 'raw_timestamp_part_1', 'raw_timestamp_part_2', 'new_window', 'num_window', 'roll_belt', 'pitch_belt', 'yaw_belt', 'total_accel_belt', 'gyros_belt_x', 'gyros_belt_y', 'gyros_belt_z', 'accel_belt_x' ,'accel_belt_y' ,'accel_belt_z' ,'magnet_belt_x' ,'magnet_belt_y' ,'magnet_belt_z' ,'roll_arm' ,'pitch_arm' ,'yaw_arm', 'total_accel_arm', 'gyros_arm_x', 'gyros_arm_y', 'gyros_arm_z', 'accel_arm_x', 'accel_arm_y', 'accel_arm_z', 'magnet_arm_x', 'magnet_arm_y', 'magnet_arm_z', 'roll_dumbbell', 'pitch_dumbbell', 'yaw_dumbbell')]
answer <- predict(finalModel, data.test.reduced)
pml_write_files(answer)
```

##References

